Efficient ancestry and mutation simulation with msprime 1.0

### 2021-11-17 response-to-reviewers.tex

\begin{reply}
We agree, and have revised Figure 2 and the caption to clarify the distinction between
stored and derived data.
\end{reply}

\begin{point}
Figure 4
It is not clear if these two plots present all the data used for the quadratic
fits, or if these data continue outside of the plots. Samples are described,
but not their ploidy, are they haploids or diploid? From the plots here, it is
not clear if the quadratic relationship holds for sample sizes << 1000. I
suggest this figure should also include a smaller sample size, (eg n=10)
\end{point}

\begin{reply}
We have moved Figure 4 to the appendix, as it is of more specialised interest
than the other main text figures. We have truncated the x-axis on the
quadratic fits to just the available data, to clarify that we are showing
all of the data. We have clarified that samples are diploid.

The plot is already quite complex and it is difficult to
see how we could add a third line without major changes. The figure is
primarily intended as a useful yardstick for users who wish to run simulations
and would like a rough idea of how long it should take, and therefore adding
a third line for very small sample sizes is likely to make it less useful
for this purpose.
\end{reply}

@@

\begin{point}
Figure 8
Averaged points are referenced but not shown "Each point...". The plot could be
more clear where the data points are and where the lines are interpolating
between data points. Because of how the simulations were described and plotted,
I found it awkward to try to compare the running time of the DTWF and
coalescent models, even if this was not the purpose of the current figure. The
figure says "to ensure we are measuring" I would suggest changing the text to
"with a goal of measuring".
\end{point}

\begin{reply}
We have update this figure (and all others in the main text) to include
data markers. We have also rephrased the caption to try to clarify the
parameters simulated and make them consistent with other figures.
\end{reply}

\begin{point}
Individual line comments
Line 142 - what is the "msp" program? Is this a command line version?
\end{point}

\begin{reply}
We have clarified this point with a parenthetical comment.
\end{reply}

\begin{point}
Line 370 - the text here makes it sound like Kelleher et al (2016) makes a
statement relating eq. 1, but the Hein et al (2004) is not cited in the Keleher
et al (2016).
\end{point}

\begin{reply}
We have rephrased this to say "(see also Kelleher et al., 2016, Fig. 2)".
\end{reply}

### 2021-11-18 paper.tex

\begin{figure}
\begin{center}
\includegraphics{figures/dtwf-perf}
\end{center}
\caption{\label{fig-dtwf-perf} Comparison of Discrete Time Wright-Fisher
(DTWF) simulation performance in \msprime\ and \ARGON\ (Intel Xeon E5-2680 CPU).
We ran simulations with a population size of $10^4$ and
recombination rate of $10^{-8}$, with 500 diploid samples
and varying sequence length.
We report (A) total CPU time and (B) maximum memory usage;
each point is the average over $5$ replicate simulations.
We show observations for \ARGON, \msprime's DTWF implementation (``DTWF'')
and a hybrid simulation of 100 generations of the DTWF followed by
the standard coalescent with recombination (``DTWF + Hudson'').
We ran \ARGON\ with a mutation rate of $0$ and with minimum output options,
with a goal of measuring only ancestry simulation time.
Memory usage for \msprime's DTWF and hybrid simulations are very similar.
}
\end{figure}

>>>>

\begin{figure}
\begin{center}
\includegraphics{figures/dtwf-perf}
\end{center}
\caption{\label{fig-dtwf-perf} Comparison of Discrete Time Wright-Fisher
(DTWF) simulation performance in \msprime\ and \ARGON\ (Intel Xeon E5-2680 CPU).
We simulated ancestry for a sample of 1000 haploids from a population of 10000,
and report the (A) total CPU time and (B) maximum memory usage for varying
sequence lengths, and a per-base recombination rate of $10^{-8}$. Each
point is the average over $5$ replicate simulations.
We show observations for \ARGON, \msprime's DTWF implementation (``DTWF'')
and a hybrid simulation of 100 generations of the DTWF followed by
the standard coalescent with recombination model (``DTWF + Hudson'').
Memory usage for \msprime's DTWF and hybrid simulations are very similar.
We ran \ARGON\ with a mutation rate of $0$ and with minimum output options,
to ensure we are measuring only ancestry simulation time.}
\end{figure}
